Evaluating Cultural Transmission in Bronze Age burial rites of Central, Northern and North-western Europe using radiocarbon data

### 2019-03-28 review/answer_to_review.md

After careful deliberation, I have decided to follow this advice. The simulation (1) does not add information that would be absolutely essential for the core topics, (2) reduces the overall comprehensibility due to the terminology that has to be introduced to describe it and (3) cannot be adequately presented and discussed here without leaving the initial scope of the paper. I will prepare another paper in the future that will present this simulation and other simulation approaches.

> Secondly, I feel like the empirical analyses have too much preprocessing. The raw data provides spatial and temporal coordinates of individual burials, yet these are combined spatially and chronologically first and analysed as aggregates. As a consequence: 1) there is a substantial loss of spatial information; 2) sampling error is effectively disregarded (i.e. mantel tests are based on proportion values that are calculated for regions with very different samples sizes) and strictly speaking the samples are also non-independent (i.e. a large site with multiple burials can strongly bias the proportions of a particular region); and 3) there is an increased risk of ecological fallacy/Simpson’s paradox. I take this aggregation provides a direct mapping to existing literature, but I am not convinced that this is worth sacrificing so much.

The decision to create discrete spatial area classes is necessary to establish subsets of dates that can be used for time series calculation. I'm not aware of a methodologival approach where time series could be constructed without some sort of aggregation. Unfortunately the size of the areas has to be very big because of the currently overall small amount of data. I expect this to change in the future and the analysis can be redone then with more data and much smaller areas to understand the definitely much more complex processes within the macro regions established here.

So: The criticism listed here is valid and I added it in the discussion. I considered a complete revision of the time series construction methodology, but since I (1) am still convinced that it is scientifically sound and reproducible, and (2) do not know a fundamentally better approach, I suggest publishing the results as they are for the community of practice to discuss and improve.

### 2019-03-28 article/04_Discussion.Rmd
further discussion of data quality

1. The amount, distribution and quality of the data as discussed in section \ref{data-quality-discussion} question the credibility of the cultural distance calculation. The macro regions have to be very big due to a lack of data and don't allow for a more precise picture.

>>>>

1. The amount, distribution and quality of the data as discussed in section \ref{data-quality-discussion} question the credibility of the cultural distance calculation. The macro regions have to be very big due to a lack of data and don't allow for a more precise picture. The differences in the number of available dates -- sample size -- per region introduce sampling error. The samples are also non-independent, as large sites with multiple burials can strongly bias the proportions of a particular region. Furthermore the region size also causes strong risk of Simpson’s paradox (i.e. a trend appears in several different groups of data but disappears or reverses when these groups are combined).
